---
title: "Appendix C: IRT Modelling with Cleaned Data"
author: "Tan, L. Y."
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    keep_md: true
    
---
<style type = "text/css">
  .main-container{max-width: 80%;}
  .row{display: flex;}
  .column{flex: 50%;}

pre.verbatim code span {
  color: inherit !important;
  font-weight: normal !important;
  font-style: normal !important;
}
</style>

```{r setup, include = FALSE, warning = FALSE}

library(dplyr)
library(ggplot2)
library(igraph)
library(readxl)
library(reshape2)
library(rmarkdown)
library(bigIRT)
library(data.table)


#Housekeeping
rm(list = ls())

knitr::knit_engines$set(text = function(options) {
  if (isFALSE(options$echo)) return("")
  paste0(
    "\n```{.r .verbatim}\n",
    paste(options$code, collapse = "\n"),
    "\n```\n"
  )
})
```


## 1. Reshaping Dataframe

The following are the codes for further preprocessing prior to applying bigIRT:

```{r format-mclean-data, eval=TRUE, warning=FALSE}

# Load the file
f <- "df_irt_itemv.csv"
hdr <- names(fread(f, nrows = 0))          # person IDs are the header


# Wide format
wide <- fread(f, header = TRUE,
              colClasses = c("character", rep("integer", length(hdr) - 1L)),
              na.strings = c("NA", "", ".", "NULL"))
setnames(wide, 1, "Item")

setDT(wide)
setnames(wide, c("Item", as.character(unlist(wide[1, -1]))))
wide <- wide[-1]
for (j in names(wide)[-1]) set(wide, j = j, value = as.integer(wide[[j]]))


# Apply melt in batches
id_col      <- "Item"
person_cols <- setdiff(names(wide), id_col)

chunks <- split(person_cols, ceiling(seq_along(person_cols) / 2000L))
out    <- vector("list", length(chunks))

for (k in seq_along(chunks)) {
  sub <- wide[, c(id_col, chunks[[k]]), with = FALSE]
  out[[k]] <- melt(sub, id.vars = id_col,
                   variable.name = "id", value.name = "score",
                   na.rm = TRUE, variable.factor = FALSE)
  cat(k, "/", length(chunks), "- kept", nrow(out[[k]]), "\n")
}

# Convert into long format for bigIRT
long <- rbindlist(out); rm(out, wide); gc()

long[, Scale := "1"]                        # Required for fitIRT()
long[, score := as.integer(score)]
setcolorder(long, c("id", "Item", "Scale", "score"))
stopifnot(!anyNA(long), all(long$score %in% c(0L, 1L)))

# Check stats before fitting
Np <- uniqueN(long$id); Ni <- uniqueN(long$Item)
#nrow(long) / (Np * Ni)                      # observed density = 0.01696261
#mean(long$score)                            # overall p-value = 0.7075237

```
<br>

### 2. Summary Statistics

#### 2.1 Per-Person Summary
Below are the summary statistics on the number of items responded per individual:

```{r per-person-summary, eval=TRUE}

# Check distribution again for person & item
per_person <- long[, .N, by = id][order(N)]

summary(per_person$N) # summary stats
sd(per_person$N)      # SD
```
<br>

And here is the per person boxplot:

```{r per-person-boxplots, eval=TRUE}
# Plot boxplots
boxplot(per_person$N)

```
<br>

#### 2.2 Per-Item Summary

Likewise, the summary statistics per item:

```{r per-item-summary, eval=TRUE}

# Check distribution again for item
per_item   <- long[, .N, by = Item][order(N)]

summary(per_item$N)   #summary stats
sd(per_item$N)        #SD

```
<br>

Per-item boxplot:

```{r boxplots, eval=TRUE}

# Plot boxplots
boxplot(per_item$N)

```
<br>

### 3. Final Cleaning Round

#### 3.1 Fragmentation checks

One last step checking for fragmentation of observations before IRT modelling (i.e., due to a lack of linking items, causing responses to form isolated groups):

```{r sparsity-checks, eval=TRUE}

# Check sparsity by analyzing the data as a graph
g <- graph_from_data_frame(
       long[, .(from = paste0("P_", id), to = paste0("I_", Item))],
       directed = FALSE)
cmp <- components(g)
cmp$no                    # want 1
sort(cmp$csize, decreasing = TRUE)[1:10]

# Keep only the largest component if fragmented
gc_v  <- names(cmp$membership)[cmp$membership == which.max(cmp$csize)]
keepP <- sub("^P_", "", grep("^P_", gc_v, value = TRUE))  # N = 45208 is the largest component
keepI <- sub("^I_", "", grep("^I_", gc_v, value = TRUE))
long  <- long[id %in% keepP & Item %in% keepI]
rm(g, cmp); gc()

```

The data formed 2 fragments, and the largesr component had N = 45208 responses. 246 responses had to be dropped due to fragmentation.

### 3.2 Removing perfect scores

Final round of cleaning to remove items / persons with perfect scores:

```{r}
# Check for perfect scores for removal
d0 <- copy(long)
d1 <- dropPerfectScores(d0, scoreref. = "score", itemref. = "Item", idref. = "id")
uniqueN(long$id)   - uniqueN(d1$id)      # persons lost
uniqueN(long$Item) - uniqueN(d1$Item)    # items lost
```

Another 1615 responses and 7 items dropped from the dataframe. Next, to fit 1-IRT using bigIRT on the final cleaned dataframe.

## 4. Fitting 1-IRT using bigIRT

Model fitting using bigIRT:

```{r big-IRT, eval=TRUE}

# Fit the model with a subset of data
set.seed(1)

fit <- fitIRT(
  long,
  score = 'score', id = 'id', item = 'Item', scale = 'Scale',
  pl    = 1,                 # Rasch: A=1, C=0, D=1 fixed
  cores = 6,                 # parallelised by splitting on id
  dropPerfectScores = TRUE,
  ebayes = TRUE,             # priors adapted from a first-pass estimate
  priors = TRUE,
  normalise = TRUE,          
  iter = 2000
)

# Extract results
itemPars   <- as.data.table(fit$itemPars)    # Item, A, B, C, D  (A=1, C=0, D=1 at pl=1)
personPars <- as.data.table(fit$personPars)  # Not important but just in case
                                             # Only had person estimates for N = 40587
```
<br>

Check which items were dropped:

```{r items-dropped, eval=TRUE}

# List items dropped
setdiff(unique(long$Item), itemPars$Item)

```
<br>

Extract and save item parameters:

```{r extraction, eval=TRUE, echo=TRUE}

# Normalize the parameters
np <- normaliseIRT(B = itemPars$B, Ability = personPars[[2]],
                   A = itemPars$A, normbase = 'Ability', robust = FALSE)


# Extract estimates
itemPars <- as.data.table(fit$itemPars)
setnames(itemPars, 1, "Item")
itemPars[, .(nItems = .N, A = uniqueN(A), C = uniqueN(C), D = uniqueN(D))]

personPars <- as.data.table(fit$personPars)   # Need this for Wright Map later
setnames(personPars, 1, "id")
theta_col <- names(personPars)[2]

```
<br>

Check correlation between item estimates and item scores:

```{r correl-items, eval=TRUE}

# Check correlation between item estimates and item scores:
pvals <- long[, .(p_correct = mean(score), n_obs = .N), by = Item]
chk   <- merge(itemPars, pvals, by = "Item")

cor(chk$B, chk$p_correct, method = "spearman")        # Should be r = -.95 or better; r = -0.9543101

```

Correlation at r = -0.95 or better is a good sign that bigIRT worked properly with the observations.

<br>

### 4.1 Summary of Resulting Item Estimates

Overall statistics of item estimates; bigIRT does not have the same functionality as WinSteps but it gives a rough idea on item separation:

```{r item-summary, eval=TRUE}

# View summary of item estimates
summary(itemPars$B)
hist(itemPars$B)

```
<br>

Plot Wright map to see how well-targeted are the person and item measures:

```{r plot-WM, eval=TRUE}
# Plot Wright map
wm <- rbind(
  data.table(logit = personPars[[theta_col]], type = "Persons (ability θ)"),
  data.table(logit = itemPars$B,              type = "Items (difficulty B)")
)
ggplot(wm, aes(logit)) +
  geom_histogram(bins = 80, fill = "grey25") +
  facet_wrap(~type, ncol = 1, scales = "free_y") +
  labs(x = "Logits", y = "Count", title = "Person–item (Wright) map") +
  theme_bw()
```
<br>

Results suggest that the items are well-targeted although there might be a floor effect.

### 4.2 Item Fit Statistics

The following are the codes for computing item fit statistics:

```{r item-fit, eval=TRUE}

# Construct data table
d <- data.table(
  item_i = as.integer(fit$dat$item),
  id_i   = as.integer(fit$dat$id),
  score  = as.integer(fit$dat$score)
)
d[, `:=`(Item  = itemPars$Item[item_i],
         id    = personPars$id[id_i],
         theta = personPars[[theta_col]][id_i],
         A = itemPars$A[item_i], B = itemPars$B[item_i],
         C = itemPars$C[item_i], D = itemPars$D[item_i])]

d[, P := C + (D - C) / (1 + exp(-A * (theta - B)))]   # 4PL; reduces to 1/(1+exp(-(θ-B))) at pl=1


d[, `:=`(W = P * (1 - P), resid = score - P)]
d[, Ck := W * ((1 - P)^3 + P^3)]        # E[(x - P)^4] for a Bernoulli

itemfit <- d[, .(
  N      = .N,
  pObs   = mean(score),
  pExp   = mean(P),
  SE_B   = 1 / sqrt(sum(W)),                    # Fisher information for B, θ treated as known
  outfit = mean(resid^2 / W),                   # unweighted mean square
  infit  = sum(resid^2) / sum(W),               # information-weighted mean square
  q_out  = sqrt(sum(Ck / W^2 - 1)) / .N,
  q_in   = sqrt(sum(Ck - W^2)) / sum(W)
), by = .(item_i, Item)]

# Wilson–Hilferty standardisation (Wright & Masters)
itemfit[, `:=`(outfit_z = (outfit^(1/3) - 1) * (3/q_out) + q_out/3,
               infit_z  = (infit^(1/3)  - 1) * (3/q_in)  + q_in/3)]

itemfit <- merge(itemfit, itemPars, by = "Item")
setorder(itemfit, B)
itemfit[]

```
<br>

Dot-and-Whisker plot for item estimates:

```{r dot-whisker-plot, eval=TRUE, warning=FALSE}

# Dot and whisker plot with CI for estimates
cp <- rbind(head(itemfit, 35), itemfit[sample(.N, 35)], tail(itemfit, 25))
cp <- unique(cp, by = "Item")                      # drop the collided rows
cp[, Item := factor(Item, levels = Item[order(B)])]

ggplot(cp, aes(B, Item)) +
      geom_errorbarh(aes(xmin = B - 1.96*SE_B, xmax = B + 1.96*SE_B), height = 0) +
      geom_point(size = 1) +
      labs(x = "Difficulty B (±1.96 SE)", y = NULL) + theme_bw()

```
<br>

Check item separation statistics:

```{r item-separation, eval=TRUE}
# Computation for item separation
itemfit[, `:=`(Wbar = 1/(SE_B^2 * N),
               d_floor = log10(SE_B) - log10(2/sqrt(N)))]

MSE  <- mean(itemfit$SE_B^2)
SDo  <- sd(itemfit$B)
G    <- sqrt(max(SDo^2 - MSE, 0)/MSE)   # item separation index
c(separation = G, reliability = G^2/(1 + G^2))
```
Results show separation at 6.25 levels with item reliability of 0.975.

<br>
<br>
<br>
<br>